---
title: "unemployment"
output: html_document
---

```{r setup, include=FALSE}
library(zoo)
library(plotly)
library(ggplot2)
library(tseries)
```

```{r}
unemploy_data <- read.csv('unemployment_data_us.csv')
unemploy_data_complete <- na.omit(unemploy_data)
```

```{r}
unemploy_data_complete$Date <- as.Date(as.yearmon(unemploy_data_complete$Date, "%b-%Y"))
rownames(unemploy_data_complete) <- 1:nrow(unemploy_data_complete)
unemploy_data_complete <- unemploy_data_complete[-c(22, 33),]
```

```{r}
fig <- plot_ly()

fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Primary_School, name = "Primary School", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~High_School, name = "High School", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Associates_Degree, name = "Associates_Degree", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Professional_Degree, name = "Professional_Degree", type = "scatter") %>% layout(title = "Unemployment Rate by Education Level")

fig
```

The plot expresses a downward trend for the past decade for the unemployment rate for all education levels.

```{r}
fig <- plot_ly()

fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~White, name = "White", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Black, name = "Black", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Asian, name = "Asian", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Hispanic, name = "Hispanic", type = "scatter") %>% layout(title = "Unemployment Rate by Race")

fig
```

The plot expresses a downward trend for the past decade for the unemployment rate for races.

```{r}
fig <- plot_ly()

fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Men, name = "Men", type = "scatter")
fig <- fig %>% add_trace(data = unemploy_data_complete, x = ~Date, y = ~Women, name = "Women", type = "scatter")  %>% layout(title = "Unemployment Rate by Gender")

fig
```

The plot expresses a downward trend for the past decade for the unemployment rate for both genders.

```{r}
unemploy_data_ts = subset(unemploy_data_complete, select = -c(Year, Month))
unemploy_data_ts <- unemploy_data_ts[order(unemploy_data_ts$Date), ]
rownames(unemploy_data_ts) <- 1:nrow(unemploy_data_ts)

Primary_School_ts = ts(unemploy_data_ts$Primary_School, start = c(2010, 1), frequency = 12)
High_School_ts = ts(unemploy_data_ts$High_School, start = c(2010, 1), frequency = 12)
Associates_Degree_ts = ts(unemploy_data_ts$Associates_Degree, start = c(2010, 1), frequency = 12)
Professional_ts = ts(unemploy_data_ts$Professional_Degree, start = c(2010, 1), frequency = 12)
White_ts = ts(unemploy_data_ts$White, start = c(2010, 1), frequency = 12)
Black_ts = ts(unemploy_data_ts$Black, start = c(2010, 1), frequency = 12)
Asian_ts = ts(unemploy_data_ts$Asian, start = c(2010, 1), frequency = 12)
Hispanic_ts = ts(unemploy_data_ts$Hispanic, start = c(2010, 1), frequency = 12)
Men_ts = ts(unemploy_data_ts$Men, start = c(2010, 1), frequency = 12)
Women_ts = ts(unemploy_data_ts$Women, start = c(2010, 1), frequency = 12)
```

```{r}
decomp<-decompose(Primary_School_ts)
plot(decomp)
decomp<-decompose(High_School_ts)  
plot(decomp) 
decomp<-decompose(Associates_Degree_ts)  
plot(decomp) 
decomp<-decompose(Professional_ts)  
plot(decomp) 
decomp<-decompose(White_ts)  
plot(decomp) 
decomp<-decompose(Black_ts)  
plot(decomp) 
decomp<-decompose(Asian_ts)  
plot(decomp) 
decomp<-decompose(Hispanic_ts)  
plot(decomp) 
decomp<-decompose(Men_ts)  
plot(decomp) 
decomp<-decompose(Women_ts)  
plot(decomp) 
```

All data shows a downward trend.
There is a seasonal support for all data happens around mid point of the year.
The random plot exhibits white noise so the best forecast is the overall mean of the historical observations.
This is additive since the seasonal changes do not enlarge over the time.

```{r}
decomp<-gglagplot(Primary_School_ts)
plot(decomp)
decomp<-gglagplot(High_School_ts)  
plot(decomp) 
decomp<-gglagplot(Associates_Degree_ts)  
plot(decomp) 
decomp<-gglagplot(Professional_ts)  
plot(decomp) 
decomp<-gglagplot(White_ts)  
plot(decomp) 
decomp<-gglagplot(Black_ts)  
plot(decomp) 
decomp<-gglagplot(Asian_ts)  
plot(decomp) 
decomp<-gglagplot(Hispanic_ts)  
plot(decomp) 
decomp<-gglagplot(Men_ts)  
plot(decomp) 
decomp<-gglagplot(Women_ts)  
plot(decomp) 
```

The positive relationship is very strong across all plots besides a bit fluctuates in Asian one, that they reflect strong seasonality in the data

```{r}
fit = lm(Primary_School_ts~time(Primary_School_ts), na.action=NULL) 
plot1 <- ggAcf(Primary_School_ts, 48, main="Original Data: Primary School")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Primary_School_ts), 48, main="first differenced data")
plot4 <- ggPacf(Primary_School_ts, 48, main="Original Data: Primary School")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Primary_School_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(High_School_ts~time(High_School_ts), na.action=NULL) 
plot1 <- ggAcf(High_School_ts, 48, main="Original Data: High School")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(High_School_ts), 48, main="first differenced data")
plot4 <- ggPacf(High_School_ts, 48, main="Original Data: High School")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(High_School_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Associates_Degree_ts~time(Associates_Degree_ts), na.action=NULL) 
plot1 <- ggAcf(Associates_Degree_ts, 48, main="Original Data: Associates Degree")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Associates_Degree_ts), 48, main="first differenced data")
plot4 <- ggPacf(Associates_Degree_ts, 48, main="Original Data: Associates Degree")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Associates_Degree_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Professional_ts~time(Professional_ts), na.action=NULL) 
plot1 <- ggAcf(Professional_ts, 48, main="Original Data: Professional")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Professional_ts), 48, main="first differenced data")
plot4 <- ggPacf(Professional_ts, 48, main="Original Data: Professional_ts")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Professional_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(White_ts~time(White_ts), na.action=NULL) 
plot1 <- ggAcf(White_ts, 48, main="Original Data: White")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(White_ts), 48, main="first differenced data")
plot4 <- ggPacf(White_ts, 48, main="Original Data: White")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(White_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Black_ts~time(Black_ts), na.action=NULL) 
plot1 <- ggAcf(Black_ts, 48, main="Original Data: Black")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Black_ts), 48, main="first differenced data")
plot4 <- ggPacf(Black_ts, 48, main="Original Data: Black")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Black_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Asian_ts~time(Asian_ts), na.action=NULL) 
plot1 <- ggAcf(Asian_ts, 48, main="Original Data: Asian")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Asian_ts), 48, main="first differenced data")
plot4 <- ggPacf(Asian_ts, 48, main="Original Data: Asian")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Asian_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Hispanic_ts~time(Hispanic_ts), na.action=NULL) 
plot1 <- ggAcf(Hispanic_ts, 48, main="Original Data: Hispanic")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Hispanic_ts), 48, main="first differenced data")
plot4 <- ggPacf(Hispanic_ts, 48, main="Original Data: Hispanic")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Hispanic_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Men_ts~time(Men_ts), na.action=NULL) 
plot1 <- ggAcf(Men_ts, 48, main="Original Data: Men")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Men_ts), 48, main="first differenced data")
plot4 <- ggPacf(Men_ts, 48, main="Original Data: Men")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Men_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)

fit = lm(Women_ts~time(Women_ts), na.action=NULL) 
plot1 <- ggAcf(Women_ts, 48, main="Original Data: Women")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Women_ts), 48, main="first differenced data")
plot4 <- ggPacf(Women_ts, 48, main="Original Data: Women")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Women_ts), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)
```

Throughout the plots, the data exhibits decaying correlation indicating that the current value of the time series is not significantly related to its past values.

In addition, the none of the plots has all the ACF and PACF inside the threshold, so they are highly possible not stationary.

```{r}
adf.test(Primary_School_ts)
adf.test(High_School_ts)
adf.test(Associates_Degree_ts)
adf.test(Professional_ts)
adf.test(White_ts)
adf.test(Black_ts)
adf.test(Asian_ts)
adf.test(Hispanic_ts)
adf.test(Men_ts)
adf.test(Women_ts)
```

Yes, the answer supports my conclusion from (5)

```{r}
Asian_ts_log <- log(Asian_ts)
fit = lm(Asian_ts_log~time(Asian_ts_log), na.action=NULL) 
plot1 <- ggAcf(Asian_ts_log, 48, main="Original Data: Asian")
plot2 <- ggAcf(resid(fit), 48, main="detrended data") 
plot3 <- ggAcf(diff(Asian_ts_log), 48, main="first differenced data")
plot4 <- ggPacf(Asian_ts_log, 48, main="Original Data: Asian")
plot5 <- ggPacf(resid(fit), 48, main="detrended data") 
plot6 <- ggPacf(diff(Asian_ts_log), 48, main="first differenced data")

grid.arrange(plot1, plot2, plot3, plot4, plot5, plot6, ncol=3, nrow=2)
```

```{r}
adf.test(Asian_ts_log)
```

I was not able to turn non-stationary into stationary for all time series data, however, for Asian, i was able to do so, the Augmented Dickey-Fuller Test with log transformation of Asian time series recorded a null hypothesis below p = 0.05 so we reject the null hypothesis and accept the alternative hypothesis that Asian in log mode is stationary.